---
title: "Introduction to Analyzing Field Data"
author: "Whalen Dillon"
date: "3/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

Install and load the package "tidyverse", which contains tools for manipulating and visualizing data. *For curriculum development, I've commented out the installation.* 

```{r}
#install.packages("tidyverse")
library("tidyverse")
```

## Import data

During lunch, we entered all of your data. Import the field data from the "data" folder.
*For curriculum development, I'm importing actual height data and simulating the rest* 

```{r read in data}

bivens_veg_data <- read_csv("coding_curriculum/data/bivens_veg_data_example.csv")
## Pooled vegetation height, weight, and moisture measured at three locations in each plot. The heights are actually averages of three measurements at each location. 
## The percent moisture is calculated as the difference between the fresh and dry weight. 

bivens_species_data <- read_csv("coding_curriculum/data/bivens_species_summer.csv")
## Reshaped (from wide to long) and subset cover data that Dr. Cathy Fahey collected. There are a lot of zeroes because it is cover data, so that is something that could be worth discussing (real data) or excluding for the exercise. 
## I simulated the 'leaves_damaged' variable

pollinators <- read_csv("coding_curriculum/data/bivens_pollinators_sim.csv")
## Simulated pollinator counts for each plot (40 values)

```


## Visualize vegetation data

```{r defining color values}

## Define discrete color and fill variables

### Manual option
# colors() ## all built-in color names
# veg_trt_color1 <- scale_color_manual(values = c("cyan","red"))
# veg_trt_fill1 <- scale_fill_manual(values = c("cyan","red"))

### Viridis option
# library(viridisLite)
# library(viridis)
# veg_trt_color2 <- scale_color_viridis(begin = .15, end = .9, option = "E", discrete = T)
# veg_trt_fill2 <- scale_fill_viridis(begin = .15, end = .9, option = "E", discrete = T)

```

### Vegetation mass
```{r fuel load fig}

ggplot(bivens_veg_data, aes(water_trt, dry_weight_g)) +
  
  geom_boxplot(aes(color = veg_trt), notch = F) +
  geom_point(aes(color = veg_trt), shape = 1, position = position_jitterdodge(jitter.width = .2)) +

  # stat_summary(aes(color = veg_trt),fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2)) +
  
  # invasion_fill +
  # invasion_color +
  theme_bw() +
  xlab("") +
  ylab(expression(paste("Vegetation mass (g ", {m}^-2, ")"), sep = "")) +
  theme(legend.position = "top")

```

```{r make a table}
knitr::kable(bivens_veg_data %>% 
    group_by(water_trt, veg_trt) %>% 
    summarise(avg_dry_wt = mean(dry_weight_g),
              nobs = length(dry_weight_g),
              se_drywt = sd(dry_weight_g)/sqrt(nobs)),
    digits = 2, caption = "Fuel load based on sample dry weight")

```


### Vegetation moisture

```{r veg moisture boxplot}

ggplot(bivens_veg_data, aes(water_trt, veg_moisture_pct, color = veg_trt)) +
  
  geom_boxplot(notch = F) +
  geom_point(shape = 1, position = position_jitterdodge(jitter.width = .2)) +
  
  # stat_summary(aes(color = veg_trt),fun.data = "mean_se", geom = "pointrange", position = position_dodge(width = .2), size = 1) +
  
  # invasion_fill +
  # invasion_color +
  theme_bw() +
  xlab("") +
  ylab("Vegetation moisture (%)") +
  theme(legend.position = "top")

## Save current figure
# ggsave("figures/veg_moisture.png", height = 4, width = 5, dpi = 300)

```

```{r veg moisture scatter plot}

ggplot(bivens_veg_data, aes(dry_weight_g, veg_moisture_pct, color = veg_trt)) +
  
  geom_point(shape = 1) +
  
  # invasion_fill +
  # invasion_color +
  theme_bw() +
  xlab("") +
  ylab("Vegetation moisture (%)") +
  theme(legend.position = "top")

```

