---
title: "invasive_species_enemies_code"
author: "Amy Kendig"
date: "3/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

Install and load the packages, which contain tools for manipulating and visualizing data. *I've included the packages within tidyverse that are already on the Architecture computers* 

```{r packages}
library(ggplot2)
library(tibble)
library(tidyr)
library(readr)
library(dplyr)
library(stringr)
library(graphics)
```

## Import data

During lunch, we entered all of your data. Import the greenhouse data from the "data" folder.
*For curriculum development, I'm importing actual height data and simulating the rest* 

```{r import data}
gh <- read_csv("./data/invasive_species_example_height.csv")
```

*Simulation code (will remove)*

```{r}
library("truncnorm")
gh <- gh %>%
  mutate(
    weight = rtruncnorm(n = 18, a = 0, mean = 4, sd = 2),
    leaf_area = rtruncnorm(n = 18, a = 0, mean = 20, sd = 10)
  ) %>%
  mutate(
    damage_area = runif(18, 0, 0.5) * leaf_area
  )
```

Let's add the full species names to the data

```{r species}
gh$species_full <- recode(gh$species, Mv = "Microstegium vimineum", Ev = "Elymus virginicus", Es = "Eragrostis spectabilis", Pc = "Panicum clandestinum", Cc = "Calamagrostis canadensis", Gs = "Glyceria striata")
```

Let's create a new column called "LMA", which is the weight of a leaf divided by its area.

```{r LMA}
gh$LMA <- gh$weight/gh$leaf_area
```

Now it's your turn to create a new column. Call this one "percent_damage" and make it equal to the damage area divided by the leaf area times 100.

```{r percent damage}
gh$percent_damage <- gh$damage_area/gh$leaf_area*100
```

Let's use ggplot to look at some of the relationships among plant traits. First, let's see how LMA and height are related to one another.

```{r height-LMA plot}
ggplot(gh, aes(x = LMA, y = height)) + 
  geom_point()
```

Now let's add on to this plot. Let's color the points by which species they belong to.

```{r height-LMA plot 2}
ggplot(gh, aes(x = LMA, y = height)) + 
  geom_point(aes(color = species))
```

What conclusions can you draw? Are all of the species similar to one another? Which is the invasive species? What are its traits like relative to the others?

Now, create a similar plot with percent damage on the y-axis and LMA on the x-axis. We've created the code chunk below for you to fill in.

```{r damage-LMA plot}


```