---
title: "invasive_species_enemies_code"
author: "Amy Kendig"
date: "3/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

Install and load the packages, which contains tools for manipulating and visualizing data. *I've included the packages within tidyverse that are already on the Architecture computers* 

```{r}
library(ggplot2)
library(tibble)
library(tidyr)
library(readr)
library(dplyr)
library(stringr)
```

## Import data

During lunch, we entered all of your data. Import the greenhouse data from the "data" folder.
*For curriculum development, I'm importing actual height data and simulating the rest* 

```{r}
gh <- read_csv("../data/invasive_species_example_height.csv")
```

Simulation code (will remove)

```{r}
library("truncnorm")
gh <- gh %>%
  mutate(
    weight = rtruncnorm(n = 18, a = 0, mean = 4, sd = 2),
    leaf_area = rtruncnorm(n = 18, a = 0, mean = 20, sd = 10)
  ) %>%
  mutate(
    damage_area = runif(18, 0, 0.5) * leaf_area
  )
```

Add the full species name to the data

```{r}
gh$species_full <- recode(gh$species, Mv = "Microstegium vimineum", Ev = "Elymus virginicus", Es = "Eragrostis spectabilis", Pc = "Panicum clandestinum", Cc = "Calamagrostis canadensis", Gs = "Glyceria striata")
```

